---
title: "Practical Worksheet"
menutitle: "Practical"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
weight: 30
---
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(rmarkdown)
library(tidyverse)
library(RColorBrewer)
library(ggpubr)
WDB.new <- read_csv("data/WDB_1999.csv")
WDB_1999 <- read_csv("data/WDB_1999.csv")
```

<font size = 4> 
**Package and Data Loading**

<br>

As mentioned within the session setup, load the following packages using the `library()` function. Additionally, as we will be using a data set with large numbers, set *scipen* to 999 using the option function. 

```{r setup, message = FALSE}

  library(tidyverse)
  library(RColorBrewer)
  library(ggpubr)

  options(scipen = 999)

```

Furthermore, for the purpose of this session, we will be using data from the [World Bank Open Data](<https://data.worldbank.org/>). In particular we will be using a collection of variables from 1999, these variables were selected to provide us plenty of room to explore! 

It is included in your downloaded zip file from the accompanying [Github Repo](<>) and can be loaded using the following code: 

```{r, eval = FALSE}

WDB_1999 <- read_csv("data/WDB_1999.csv")

```

It includes the following variables: 

* Country Name (Country Name)
* Country Code (Country Code)
* Continent (Continent)
* Year (Year)
* Population (Pop)
* Female Population (Pop.fe)
* Male Population (Pop.ma)
* Birth Rate, crude per 1000 people (birthrate)
* Death Rate, crude per 1000 people (deathrate)
* Life Expectancy at Birth in years (lifeexp)
* Female Life Expectancy at Birth in years (lifeexp.fe)
* Male Life Expectancy at Birth in years (lifeexp.ma)
* Educational Spending, percentage of GDP (ed.spend)
* Compulsory Education Duration in Years (ed.years)
* Labour Force Total (labour)
* Literature Rate in adults, percentage % (lit.rate.per)
* CO2 Emissions, kt (co2)
* Gross Domestic product, $ (gdp)
* Unemployment, percentage of total labour force (unemp)
* Female Unemployment, percentage of total labour force (unemp.fe)
* Male Unemployment, percentage of total labour force (unemp.ma)
* Health Expenditure per capita, $ (health.exp)
* Hospital Beds per 1000 people (medbeds)
* Number of Surgical Procedures per 1000 people (surg.pro)
* Number of Nurses & Midwives per 1000 people (nurse.midwi)

---

