---
title: "Beginners Worksheet"
menutitle: "Beginner"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
weight: 10
---
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
  library(rmarkdown)
  library(tidyverse)
  library(RColorBrewer)
  library(ghibli)
  library(palettetown)

```

<font size = 4> 
**Package Loading**

<br>

As mentioned within the session setup, load the following packages using the `library()` function.

```{r setup, message = FALSE}

  library(tidyverse)
  library(RColorBrewer)
  library(ghibli)
  library(palettetown)

```

---

**Exercise 1: Using the data provided (`ex1.dat`), generated by the code below, plot the data onto a scatterplot using `ggplot()`. Plotting the variable `sine` from the data onto the x variable, and the index (`1:1001`) onto the y variable**

<br> 

Setup Code

```{r, eval = TRUE, message = FALSE, warning = FALSE}

# Generate the sequence 1 to 100, in steps of 0.1
  ex1.dat <- as.data.frame(
    seq(from = 0, 
        to = 100, 
        by = 0.1))

# Apply the sine function
  ex1.dat <- sin(ex1.dat)

# Rename the columns
  colnames(ex1.dat) <- "sine"

```
<br>

Template 

```{r eval = FALSE}
  ggplot(data = ??,                  ## Replace ?? with the data which should be used.
         mapping = aes(x = ??,       ## Replace ?? with the x variable (index)
                       y = ??)) +    ## Replace ?? with the y variable (sine)
    geom_point()
```

---

**Exercise 1, Bonus Question: Rather than using geom_point(), use geom_line() or another geom_ function to plot this same data in another way.**

---

**Exercise 2: Add one of the two following themes to clean up your code!**

```{r eval = FALSE}

  theme(
    panel.grid.major = element_blank(), 	
    panel.grid.minor = element_blank(),
    panel.background = element_blank(), 	
    axis.line = element_blank(), 
    axis.title = element_blank(), 
    axis.text = element_blank(),
    axis.ticks = element_blank()) 

  theme_void()

```

---

**Exercise 3: Using the same data as before (ex1.dat) transform the data and layer it onto the previous plots. Use the geom_line() function to more easily observe the impact of this transformation.**

**Hint: Use one of the following as an example to base your transformation on**

```{r eval = FALSE}
  geom_line(data = (ex1.dat + 1),
            mapping = aes(x = 1:1001, y = sine))

  geom_line(data = (ex1.dat - 1),
            mapping = aes(x = 1:1001, y = sine))
    
  geom_line(data = (ex1.dat * 2),
            mapping = aes(x = 1:1001, y = sine))
    
  geom_line(data = (ex1.dat / 2),
            mapping = aes(x = 1:1001, y = sine))
```

**Exercise 3 Bonus: Repeat this progress building multiple layers**

---

**Exercise 4: Changing Coordinate System**

<br>

**Step 1: Create a new data set ex4.dat using the following code**

```{r}
ex4.dat <- as.data.frame(seq(from = 1,
                                 to = 51.3, 
                                 by = 0.1))
    ex4.dat <- sin(ex4.dat)
    colnames(ex4.dat) <- "sine" 

```

**Step 2: Using the code we have previously used, add the function coord_polar()**

**Hint: Use "mapping = aes(x = 1:504, y = sine)" for ex4.dat** 

<br>

**Step 3: Once again, repeat the transformation adding new layers to this new plot**

<br>

**Exercise 4 Bonus: Return to the sequence generation in step 1, How does this plot change if a function like cos() or tan() is applied rather than tan()**

---

**Exercise 5: Layering Colours, using the code created in Exercise 4, replace `geom_point()` with `geom_polygon()` and apply a colour within using `fill =` or `colour = `.**

<br>

To view all the colours which can be used: `display.brewer.all()`, or see: [this cheatsheet](<https://www.nceas.ucsb.edu/sites/default/files/2020-04/colorPaletteCheatsheet.pdf>)

<br>

Template

<br> 

Colours could include: 

- Blue: "blue", #0074D9
- Red: "red", #FF4136

Or from packages:

- RColorBrewer
- ghibli
- palettetown

<br>

```{r eval = FALSE}
geom_polygon(colour = XX)
```

---

**Exercise 6: Exploration and Experimentation, using the skills previously known and the other topics discussed in the learning material section, explore and experiment to see what types of patterns you can create. Look at the inspiration section for ideas.**

<br>

**Hint and tips**

- Change the core data set,
- Layer different data sets together,
- Try different geom functions 
- Try different colours, layers or palettes. 

---