---
title: "Practical Introduction"
menutitle: "Practical"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
weight: 30
---
--- 

<font size = 4>
**Package and Data Loading**

<br>

As mentioned during the session setup, load the following packages using the `library()` function. Additionally, as we will be using a data set which contains large numbers, set scipen to 999, within the `options()` function.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(caret)
library(tree)
library(rpart)
library(rpart.plot)

options(scipen = 999)
```

Furthermore, for the purpose of this session, we will be using data from [UC Irvine Machine Learning Repository](https://archive-beta.ics.uci.edu/) and in particular the [Red Wine Quality Dataset](<https://archive-beta.ics.uci.edu/ml/datasets/109>). This can be downloaded from the site directly, or contained in the [.zip file](</zip/introml_tree.zip>).

```{r, warning = FALSE, message = FALSE}
winequality_red <- read_csv("data/winequality-red.csv")
```

<br>

---

<br>

This data set contains 1599 observations across the following variables: 

* Fixed Acidity ('fixed acidity') - fixed & nonvolatile acid in the wine. 
* Volatile Acidity ('volatile acidity') - the amount of acetic acid in the wine, related to vinegar flavours.
* Citric Acid ('citric acid') - amount of citric acid in the wine, related to "freshness".
* Residual Sugar ('residual sugar') - the amount of sugar remaining after fermentation.
* Chlorides ('chlorides') - the amount of salt in the wine.
* Free Sulfur Dioxide ('free sulfur dioxide') - A free form of SO2, preventing microbial growth and oxidation. 
* Total Sulfur Dioxide ('Total Sulfur Dioxide') - Total SO2 amount, calculated from both free and bound forms. 
* Density ('density') - The density of water, depending on the percent of alcohol and sugar content.
* pH ('pH') - Describes how acidic or basic the wine is (between 0, very acidic, and 14, very basic).
* Sulphates ('sulphates') - A wine additive which contributes to SO2 levels, acts as an antimicrobial and antioxidant.
* Alcohol ('alcohol') - The percent alcohol content of the wine. 
* Quality ('quality') - The output variable, based upon sensory data (0 - 10).

<br>

--- 

<br>
