---
title: "Solutions"
menutitle: "Solutions"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
weight: 30
---
---

<font size = 4> 
**Package and Data Loading**

<br>

As mentioned within the session setup, load the following packages using the `library()` function. Additionally, as we will be using a data set with large numbers, set *scipen* to 999 using the option function. 

```{r setup, message = FALSE}

  library(tidyverse)
  library(RColorBrewer)

  options(scipen = 999)

```

Furthermore, for the purpose of this session, we will be using data from the [World Bank Open Data](<https://data.worldbank.org/>). In particular we will be using a collection of variables from 1999, these variables were selected to provide us plenty of room to explore! 

It is included in your downloaded zip file from the accompanying [Github Repo](<>) and can be loaded using the following code: 

```{r, message = FALSE, warning = FALSE}

WDB_1999 <- read_csv("../data/WDB_1999.csv")

```

---

<font size = 4>
**Section 1: ggplot2 vs plot**

<br>

**Exercise 1: Plotting *birthrate* against *deathrate* using both the `plot()` and `ggplot()` function, discuss which has more potential in displaying data clearly.**

```{r}

## Plotting using plot()
  
  plot(data = WDB_1999,
       birthrate ~ deathrate)


## Plotting using ggplot()

  ggplot(data = WDB_1999, 
         mapping = aes(
           x = deathrate,
           y = birthrate
         )) + 
    geom_point()

```

<br>

**Exercise 2: Expand the plot to group these points by Continent, which provides us with more information and is easier to achieve? Remember, you'll need to recode WDB_1999$Continent as a factor using the function:**

```{r}

## Plotting using plot()

WDB_1999$Continent <- as.factor(WDB_1999$Continent)
  
  plot(data = WDB_1999,
       birthrate ~ deathrate,
       col = c("blue", "light blue", "red", "pink", 
             "green", "dark green", "black", "purple")[Continent])


## Plotting using ggplot()

  ggplot(data = WDB_1999, 
         mapping = aes(
           x = deathrate,
           y = birthrate,
           colour = Continent
         )) +
    geom_point()

```

<br>

---

<font size = 4>
**Section 2: Scatter Plots in ggplot**

<br>

**Exercise 3: Change the size parameter to *ed.years* to see if there is a trend between amount of years in Education and the Birth and Death Rate, set the alpha parameter to 0.5 to clearly see the relationships.**

```{r}

  ggplot(data = WDB_1999, 
         mapping = aes(
           x = birthrate,
           y = deathrate,
           colour = Continent,
           size = ed.years
         )) +
    geom_point(alpha = 0.5)

```

<br>

**Exercise 4: Change the Labels on the X and Y axis' and provide a suitable title for the graph** 

```{r}

  ggplot(data = WDB_1999, 
         mapping = aes(
           x = birthrate,
           y = deathrate,
           colour = Continent,
           size = ed.years
         )) +
    geom_point(alpha = 0.5) + 
  labs(x = "Birth Rate", 
       y = "Death Rate",
       title = "Interaction between Birth and Death Rate.")

```

<br>

---

<font size = 4>
**Section 3: Bar Charts and Histograms**

<br>

**Exercise 5: Using the parameter `stat = "identity"` within the `geom_bar()` function, create a bar chart of `Continent` plotted against the mean `birthrate` or `deathrate`**

```{r}

  ggplot(data = WDB_1999) + 
    geom_bar(stat = "identity", 
             mapping = aes(x = Continent, 
                           y = mean(birthrate, na.rm = TRUE))) 

  ggplot(data = WDB_1999) + 
    geom_bar(stat = "identity", 
             mapping = aes(x = Continent, 
                           y = mean(deathrate, na.rm = TRUE))) 

```

<br>

**Exercise 6:  Using the function `geom_histogram()` create a histogram of the *birthrate* and *deathrate* ** 

```{r}

  ggplot(data = WDB_1999) + 
    geom_histogram(mapping = aes(x = birthrate), colour = "blue", alpha = 0.5) + 
    geom_histogram(mapping = aes(x = deathrate), colour = "red", alpha = 0.5)

```

<br>

---

<font size = 4>
**Section 4: Adding density plots to Histograms**

<br>

**Exercise 7: Using the plot created in exercise 6, add the y-variable `..density..` and `binwidth = 1` to `geom_histogram()` in addition to adding `geom_density()` to add density lines to the Histogram**

```{r}

  ggplot(data = WDB_1999) + 
    geom_histogram(mapping = aes(x = birthrate, y = ..density..), binwidth = 1, 
                   colour = "blue", alpha = 0.5) + 
    geom_histogram(mapping = aes(x = deathrate, y = ..density..), binwidth = 1, 
                   colour = "red", alpha = 0.5) + 
    geom_density(mapping = aes(x = birthrate), colour = "blue", alpha = 0.5) + 
    geom_density(mapping = aes(x = deathrate), colour = "red", alpha = 0.5)

```

<br>

**Exercise 8: Add the parameter, `adjust = 2` in the density plot, to smooth this link and make it more easily interpretable**

```{r, eval = FALSE}

  ggplot(data = WDB_1999) + 
    geom_histogram(mapping = aes(x = birthrate, y = ..density..), binwidth = 1, 
                   colour = "blue", alpha = 0.5) + 
    geom_histogram(mapping = aes(x = deathrate, y = ..density..), binwidth = 1, 
                   colour = "red", alpha = 0.5) + 
    geom_density(mapping = aes(x = birthrate), colour = "blue", alpha = 0.5, adjust = 2) + 
    geom_density(mapping = aes(x = deathrate), colour = "red", alpha = 0.5, adjust = 2)

```

<br>

---

<font size = 4>
**Section 5: Extra Useful Tips and Functions**

<br>

**Exercise 9: Use the `ggsave()` function to save your last plot**

```{r, eval = FALSE}

ggsave(filename = ??, 
       plot = last_plot())

```