---
title: "Practical Worksheet"
menutitle: "Practical"
mainfont: Arial
fontsize: 12pt
urlcolor: blue
weight: 30
---
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
library(rmarkdown)
library(tidyverse)
WDB.new <- read_csv("data/WDB_1999.csv")
```

<font size = 4> 
**Package and Data Loading**

<br>

As mentioned within the session setup, load the following packages using the `library()` function. Additionally, as we will be using a data set with large numbers, set *scipen* to 999 using the option function. 

```{r setup, message = FALSE}

  library(tidyverse)
  library(RColorBrewer)

  options(scipen = 999)

```

Furthermore, for the purpose of this session, we will be using data from the [World Bank Open Data](<https://data.worldbank.org/>). In particular we will be using a collection of variables from 1999, these variables were selected to provide us plenty of room to explore! 

It is included in your downloaded zip file from the accompanying [Github Repo](<>) and can be loaded using the following code: 

```{r, eval = FALSE}

WDB_1999 <- read_csv("data/WDB_1999.csv")

```

It includes the following variables: 

* Country Name (Country Name)
* Country Code (Country Code)
* Continent (Continent)
* Year (Year)
* Population (Pop)
* Female Population (Pop.fe)
* Male Population (Pop.ma)
* Birth Rate, crude per 1000 people (birthrate)
* Death Rate, crude per 1000 people (deathrate)
* Life Expectancy at Birth in years (lifeexp)
* Female Life Expectancy at Birth in years (lifeexp.fe)
* Male Life Expectancy at Birth in years (lifeexp.ma)
* Educational Spending, percentage of GDP (ed.spend)
* Compulsory Education Duration in Years (ed.years)
* Labour Force Total (labour)
* Literature Rate in adults, percentage % (lit.rate.per)
* CO2 Emissions, kt (co2)
* Gross Domestic product, $ (gdp)
* Unemployment, percentage of total labour force (unemp)
* Female Unemployment, percentage of total labour force (unemp.fe)
* Male Unemployment, percentage of total labour force (unemp.ma)
* Health Expenditure per capita, $ (health.exp)
* Hospital Beds per 1000 people (medbeds)
* Number of Surgical Procedures per 1000 people (surg.pro)
* Number of Nurses & Midwives per 1000 people (nurse.midwi)

---

<font size = 4>
**Section 1: ggplot2 vs plot**

<br>

**Exercise 1: Plotting *birthrate* against *deathrate* using both the `plot()` and `ggplot()` function, discuss which has more potential in displaying data clearly.**

<br>

**Exercise 2: Expand the plot to group these points by Continent, which provides us with more information and is easier to achieve? Remember, you'll need to recode WDB_1999$Continent as a factor using the function:**

```{r}

WDB_1999$Continent <- as.factor(WDB_1999$Continent)

```

<br>

---

<font size = 4>
**Section 2: Scatter Plots in ggplot**

<br>

**Exercise 3: Change the size parameter to *ed.years* to see if there is a trend between amount of years in Education and the Birth and Death Rate, set the alpha parameter to 0.5 to clearly see the relationships.**

<br>

**Exercise 4: Change the Labels on the X and Y axis' and provide a suitable title for the graph** 

<br>

---

<font size = 4>
**Section 3: Bar Charts and Histograms**

<br>

**Exercise 5: Using the parameter `stat = "identity"` within the `geom_bar()` function, create a bar chart of `Continent` plotted against the mean `birthrate` or `deathrate`**

<br>

**Exercise 6:  Using the function `geom_histogram()` create a histogram of the *birthrate* and *deathrate* ** 

<br>

---

<font size = 4>
**Section 4: Adding density plots to Histograms**

<br>

**Exercise 7: Using the plot created in exercise 6, add the y-variable `..density..` and `binwidth = 1` to `geom_histogram()` in addition to adding `geom_density()` to add density lines to the Histogram**

<br>

**Exercise 8: Add the parameter, `adjust = 2` in the density plot, to smooth this link and make it more easily interpretable**

<br>

---

<font size = 4>
**Section 5: Extra Useful Tips and Functions**

<br>

**Exercise 9: Add a `legends()` and `breaks()` to a plot**

<br>

**Exercise 10: Use the `ggsave()` function to save your last plot**
